---
title: "Reproducible research assignment 1"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.

The variables included in this dataset are:

* steps: Number of steps taking in a 5-minute interval (missing values are coded as NA)
* date: The date on which the measurement was taken in YYYY-MM-DD format
* interval: Identifier for the 5-minute interval in which measurement was taken

The dataset is stored in a comma-separated-value (CSV) file and there are a total of 17,568 observations in this dataset.

## Description of assignment
1. Code for reading in the dataset and/or processing the data
2. Histogram of the total number of steps taken each day
3. Mean and median number of steps taken each day
4. Time series plot of the average number of steps taken
5. The 5-minute interval that, on average, contains the maximum number of steps
6. Code to describe and show a strategy for imputing missing data
7. Histogram of the total number of steps taken each day after missing values are imputed
8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
9. All of the R code needed to reproduce the results (numbers, plots, etc.) in the report

## Loading and preprocessing the data
Load libraries
```{r load libraries, echo=FALSE}
library(dplyr)
library(lubridate)
library(ggplot2)
library(tidyr)
library(imputeTS)
```
### Read the data
```{r read data}
df0 <- read.csv(file = "./raw_data/activity.csv")
```
### Check data and field formatting
```{r check data}
head(df0)
str(df0)
summary(df0)
```
### Convert date to POSIX and remove NA
```{r convert date to POSIXct}
df <- na.omit(df0)
df[,2] <- as.POSIXct(df[,2], origin = "2012-10-01", format = "%Y-%m-%d")
```
## Calculate total number of steps per day and print histogram
```{r calculate mean per day}
df2 <- df %>%
  mutate(date = floor_date(date)) %>%
  group_by(date) %>%
  summarize(total_steps = sum(steps, na.rm = TRUE))
hist(df2$total_steps)
```
### Calculate mean and median of the total number of steps taken per day:
```{r mean and median}
print("Mean:")
mean(df2$total_steps)
print("Median:")
median(df2$total_steps)
```
### Time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r print timeseries}
df3 <- df %>%
  group_by(interval) %>%
  summarize(mean_steps = mean(steps))

p <- ggplot(df3, aes(x=interval, y=mean_steps)) +
  geom_line() + 
  labs(title = "Activity data (NAs removed)", y = "Number of steps", x = "Interval")
p
```


### Interval, on average across all the days in the dataset, that contains the maximum number of steps - 835 (average steps: 206)
``` {r max row}
df3[which.max(df3$mean_steps),]
```
## Calculate and report the total number of missing values in the dataset
```{r report on NA}
colSums(is.na(df0))

```
### Filling the missing NA using imputeTS package. Calculating necessary statistics.
```{r using tidyr to fill the missing data}
df_clean <- na_mean(df0)
#Converting date to POSIXct
df_clean[,2] <- as.POSIXct(df_clean[,2], origin = "2012-10-01", format = "%Y-%m-%d")
df_clean2 <- df_clean %>%
  mutate(date = floor_date(date)) %>%
  group_by(date) %>%
  summarize(total_steps = sum(steps, na.rm = TRUE))
hist(df_clean2$total_steps)
print("Mean:")
mean(df_clean2$total_steps)
print("Median:")
median(df_clean2$total_steps)
```
### Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day
```{r Create factor describing weekends/weekdays}
df_clean$weekd <- ifelse(weekdays(df_clean$date) %in% c("Saturday", "Sunday"), "weekend", "weekday")

```
Print data in panel plot grouped by weekday/weekend
```{r print data}
df_clean %>%
  group_by(weekd, interval) %>%
  summarize(mean_steps = mean(steps)) %>%

ggplot(aes(x=interval, y=mean_steps)) +
  geom_line() + 
  facet_wrap(~ weekd, nrow = 2) +
  labs(title = "Data with NAs filled in", y = "Number of steps", x = "Interval")
```
